---
title: "assign_inmobi"
output: html_document
---

```{r setup, include=FALSE}
d = read.table("analyst_assignment_data_2.txt", 
               sep="\t", 
               col.names=c("date", "udid","operator","state","city","zipcode","ad_requests"), 
               fill=FALSE, 
               strip.white=TRUE)
df<-d
```
Automated code for univariate analysis
```{r}
rankfreq <- function(data, column, min=1) {
  df = as.data.frame(sort(table(data[, column]), decreasing = T))
  df = df[df$Freq >= min, ]
  df$Rank = seq(length(df$Freq))
  plot(df$Rank, df$Freq, log='xy', type='b', main=column)
  df
}

ordfreq = function(data, column, levels) {
  plot(factor(data[, column], levels=levels), main=column)
  as.data.frame(table(data[, column]))
}

contfreq = function(data, column) {
  hist(data[, column], breaks='Sturges')
  as.data.frame(table(data[, column]))
}
```
Creating derived variables
```{r}
df$date<-as.Date(df$date, format = "%Y-%m-%d")
head(df)
lapply(df,class)

library(lubridate)
df$month<-month(df$date)
df$dayofweek<-weekdays(df$date)
df$weeknum<-format(df$date,"%U")
```
Rank frequency plot to understand the dataset variables
```{r}
library(dplyr)
df<-as.data.frame(tbl_df(df))

rankfreq(df,'date')
ordfreq(df,'month')
ordfreq(df,"dayofweek")
ordfreq(df,"weeknum")
```

```{r}
rankfreq(df,'operator')
rankfreq(df,'state')
rankfreq(df,'city')
rankfreq(df,'zipcode')
ordfreq(df,'ad_requests')
```
Analyze Market Share trends by Cities (Mumbai, Bangalore, Delhi, Chennai, Kolkata, Hyderabad)
```{r}
# creating dataset to analyse market share trend from April to May
market<-df %>% group_by(state,operator,date) %>% filter(city %in% c("MUMBAI","KOLKATA","BANGALORE","CHENNAI","HYDERABAD") | state %in% "DELHI" ) %>% summarise(users=n_distinct(udid))

```
To understand the churn pattern of users i.e. in which states/cities we are seeing a major movement of users to other carriers
```{r}
library(tidyr)

##can one user have more than one subscriber per month #
apr<-df[df$month==4,]
may<-df[df$month==5,]

# data for april and may in customer-operator level
check<-apr %>% select(udid,operator) %>% group_by(udid,operator)
apr_churn<-as.data.frame(check[!duplicated(check),])

check_may<-may %>% select(udid,operator) %>% group_by(udid,operator)
may_churn<-as.data.frame(check_may[!duplicated(check_may),])

#Left join on customers in Apr and May to find churn based on operator mismatch
churn_check<-merge(apr_churn,may_churn,by=c("udid"),all.x = TRUE)

# If the operator in April is not equal to May, then it is churn
churn_check$move<-ifelse(churn_check$operator.x==churn_check$operator.y,0,1)

table(churn_check$move)

# 0       1 
# 7788175  700351 

#Assumption - every operator to have a new customer even if 1 person has 2 numbers in different operators

# Percentage churn
churn_perc<-700351/(8488526)

# Loyal 
loyal_perc<-7788175/(8488526)
```
Churn at state level
```{r}
library(dplyr)

#to get distinct customers and state
cust_same_state = unique(apr[,c('date','udid','state')])
head(cust_same_state)

#Subsetting May dataset to retain only those customers who stayed in the same state
cust_may = merge(may, cust_same_state[,c('udid','state')], by= c('udid','state'))
#cust_may.head()

#Inner join at state level and then left join operators
cust_state_inner<-merge( apr[,c('date','udid','state','operator')], cust_may[,c('date','udid','state','operator')], by= c('udid','state'))

#removing duplicates
cust_state_inner<-as.data.frame(cust_state_inner[!duplicated(cust_state_inner),])

# If operator in April and May is not the same for the same user_id and state then, it is churn
cust_state_inner$check<-ifelse(cust_state_inner$operator.x==cust_state_inner$operator.y,"loyal","churn")

## checking for loyal and churn ##
table(cust_state_inner$check)

#churn   loyal 
#456163 6158790 

#state level count of churned customers
churned_state<-cust_state_inner %>% group_by(state) %>% filter(check %in% "churn") %>% summarise(udid=n_distinct(udid))

#state level count of loyal customers
loyal_state<-cust_state_inner %>% group_by(state) %>% filter(check %in% "loyal") %>% summarise(udid=n_distinct(udid))

#merging to create master output
final_churn_state<-merge(loyal_state, churned_state,by="state", suffixes=c('_loyal','_churned'))

#calculating total number of customers per state
final_churn_state$total<-final_churn_state$udid_loyal + final_churn_state$udid_churned

#calculating percentage churn per state
final_churn_state$perc_churn<-final_churn_state$udid_churned/final_churn_state$total

#calculating index per state=state churn/overall churn
final_churn_state$index<-final_churn_state$perc_churn/0.08

#higher the index, higher is the churn rate
final_churn_state<-final_churn_state[order(final_churn_state$index,decreasing = TRUE),]
head(final_churn_state)
```

```{r}
## city level churn ##
#to get distinct customers and state
cust_same_city = apr[,c('date','udid','state','city')]
cust_same_city<-as.data.frame(cust_same_city[!duplicated(cust_same_city),])

head(cust_same_city)

#Subset May dataset to retain only those customers who stayed in the same city
cust_may_city = merge(may, cust_same_city[,c('udid','state','city')], by= c('udid','state','city'))
#cust_may.head()

#Inner join at city level and then left join operators
cust_city_inner = merge( apr[,c('date','udid','state','city','operator')], cust_may_city[,c('date','udid','state','city','operator')], by= c('udid','state','city'))

cust_city_inner<-as.data.frame(cust_city_inner[!duplicated(cust_city_inner),])

cust_city_inner$check<-ifelse(cust_city_inner$operator.x==cust_city_inner$operator.y,"loyal","churn")

## checking for loyal and churn ##
table(cust_city_inner$check)

churned_city<-cust_city_inner %>% group_by(state,city) %>% filter(check %in% "churn") %>% summarise(udid=n_distinct(udid))
loyal_city<-cust_city_inner %>% group_by(state,city) %>% filter(check %in% "loyal") %>% summarise(udid=n_distinct(udid))
final_churn_city<-merge(loyal_city, churned_city,by=c("state","city"), suffixes=c('_loyal','_churned'))
final_churn_city$total<-final_churn_city$udid_loyal + final_churn_city$udid_churned

final_churn_city$perc_churn<-final_churn_city$udid_churned/final_churn_city$total
final_churn_city$index<-final_churn_city$perc_churn/0.08

final_churn_city<-final_churn_city[order(final_churn_city$index,decreasing = TRUE),]

final_churn_city
```
Top zipcodes which has seen significant movement of market share from Airtel to other operators
```{r}
library(reshape2)

# data for april in operator-zipcode-state level
apr_zipcode<-apr %>% select(operator,zipcode,state) %>% group_by(operator,zipcode,state) %>% summarise(udid=n())
apr_zipcode<-apr_zipcode[!duplicated(apr_zipcode),]

# changing from long to side for ease of analysis and calculation
apr_cast<-dcast(apr_zipcode,zipcode+state~operator,value.var = 'udid')

#calculating market share of all operators
apr_cast$airtel_share<-apr_cast$`Bharti Airtel`/rowSums(apr_cast[,c(3:6)])
apr_cast$jio_share<-apr_cast$Jio/rowSums(apr_cast[,c(3:6)])
apr_cast$idea_share<-apr_cast$Idea/rowSums(apr_cast[,c(3:6)])
apr_cast$vodafone_share<-apr_cast$`Vodafone Essar`/rowSums(apr_cast[,c(3:6)])

# data for april in operator-zipcode-state level
may_zipcode<-may %>% select(operator,zipcode,state) %>% group_by(operator,zipcode,state) %>% summarise(udid=n())

# changing from long to side for ease of analysis and calculation
may_cast<-dcast(may_zipcode,zipcode+state~operator,value.var = 'udid')

#calculating market share of all operators
may_cast$airtel_share<-may_cast$`Bharti Airtel`/rowSums(may_cast[,c(3:6)])
may_cast$jio_share<-may_cast$Jio/rowSums(may_cast[,c(3:6)])
may_cast$idea_share<-may_cast$Idea/rowSums(may_cast[,c(3:6)])
may_cast$vodafone_share<-may_cast$`Vodafone Essar`/rowSums(may_cast[,c(3:6)])

#merging april and may dataset on zipcode
zipcode_final<-merge(apr_cast[,c(1,2,7:10)],may_cast[,c(1,2,7:10)],by=c("state","zipcode"),all.x = TRUE,suffixes = c('_apr','_may'))

# calculating growth in market share for each operator per zipcode
zipcode_final$airtel_growth<-zipcode_final$airtel_share_may-zipcode_final$airtel_share_apr
zipcode_final$jio_growth<-zipcode_final$jio_share_may-zipcode_final$jio_share_apr
zipcode_final$idea_growth<-zipcode_final$idea_share_may-zipcode_final$idea_share_apr
zipcode_final$vodafone_growth<-zipcode_final$vodafone_share_may-zipcode_final$vodafone_share_apr

# sorting in ascending order to check for pincodes with maximum degrowth in Airtel
zipcode_final<-zipcode_final[order(zipcode_final$airtel_growth,decreasing = FALSE),]
zipcode_final
```
Suggest a distinctive strategy for Airtel to retain its users across 3 different states
```{r}
# Top states where airtel is experiencing churn #

# obtaining competitor churn following up on pincode analysis
# airtel_state_degrowth<-zipcode_final %>% group_by(state) %>% summarise(airtel_growth=mean(airtel_growth,na.rm=TRUE))
# airtel_state_degrowth<-airtel_state_degrowth[order(airtel_state_degrowth$airtel_growth,decreasing = FALSE),]
# airtel_state_degrowth

#obtaining data at operator-state level for april
apr_state<-apr %>% group_by(operator,state) %>% summarise(cust=n())
apr_cast<-dcast(apr_state,state~operator,value.var = 'cust')
#Market share for Airtel in April
apr_cast$airtel_share<-apr_cast$`Bharti Airtel`/rowSums(apr_cast[,c(2:5)])

#obtaining data at operator-state level for may
may_state<-may %>% group_by(operator,state) %>% summarise(cust=n())
may_cast<-dcast(may_state,state~operator,value.var = 'cust')
may_cast$airtel_share<-may_cast$`Bharti Airtel`/rowSums(may_cast[,c(2:5)])

# Left join on state
state_final<-merge(apr_cast[,c(1,6)],may_cast[,c(1,6)],by=c("state"),all.x = TRUE,suffixes = c('_apr','_may'))

# Airtel market share growth percentage
state_final$airtel_growth<-state_final$airtel_share_may-state_final$airtel_share_apr

state_final<-state_final[order(state_final$airtel_growth,decreasing = FALSE),]
state_final
```

Alternative approach to finiding the top 10 pincodes with significant movement of Market share
```{r}
# ANOTHER APPROACH TO PINCODE #

## april : zipcode - airtel + other
## may : zipcode - airtel + other

# data at operator-zipcode-customer-state level
pincode_apr<-apr %>% select(operator,zipcode,udid,state) %>% group_by(operator,zipcode,udid,state)
pincode_apr<-as.data.frame(pincode_apr[!duplicated(pincode_apr),])
pin_apr<-pincode_apr
pin_apr$op<-ifelse(!(pincode_apr$operator %in% "Bharti Airtel"),"others","airtel")
pin_apr<-pin_apr %>% group_by(op,zipcode,state) %>% summarise(udid=n())
pin_apr<-dcast(pin_apr,zipcode+state~op,value.var='udid')

#finding Market share of Airtel in April at a zipcode level
pin_apr$apr_share<-pin_apr$airtel/(pin_apr$airtel+pin_apr$others)

pincode_may<-may %>% select(operator,zipcode,udid,state) %>% group_by(operator,zipcode,udid,state)

# Inner join on zipcode-state-customerid to obtain operator movement from April to May
pin_master<-merge(pincode_apr,pincode_may,by=c("zipcode","state","udid"),suffixes = c('_apr','_may'))

#Airtel to Airtel movement
pin_master$airteltoairtel<-ifelse(pin_master$operator_apr=="Bharti Airtel" & pin_master$operator_may=="Bharti Airtel",1,0)
# Airtel to Others movement
pin_master$airteltoother<-ifelse(pin_master$operator_apr=="Bharti Airtel" & pin_master$operator_may!="Bharti Airtel",1,0)
# Other operators to Airtel movement
pin_master$othertoairtel<-ifelse(pin_master$operator_apr!="Bharti Airtel" & pin_master$operator_may=="Bharti Airtel",1,0)
# Others to others movement
pin_master$othertoother<-ifelse(pin_master$operator_apr!="Bharti Airtel" & pin_master$operator_may!="Bharti Airtel",1,0)
# Others to Jio movement
pin_master$alltojio<-ifelse(pin_master$operator_apr!="Jio" & pin_master$operator_may=="Jio",1,0)

pin_overall<-pin_master %>% group_by(zipcode,state) %>% summarise(airteltoairtel=sum(airteltoairtel),airteltoother=sum(airteltoother),othertoairtel=sum(othertoairtel),othertoother=sum(othertoother),alltojio=sum(alltojio),udid=n_distinct(udid))

# May market share = Airtel-Airtel - Airtel-Others + Others-Airtel
pin_overall$may_share<-(pin_overall$airteltoairtel-pin_overall$airteltoother+pin_overall$othertoairtel)/rowSums(pin_overall[,c(3:6)])

#Degrowth
pin_final$degrowth<-pin_final$apr_share-pin_final$may_share
```

